Hydrodynamics of a Semipermeable Inextensible Membrane Under Flow and Confinement

### 2021-02-15 vesicle_code/docs/permeableVes/semipermeable_vesicle.tex
edited up to start of section V

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a) Snapshots of four
  semipermeable inextensible membranes with $\beta = 10^{-8}$ submerged in a fluid
  with no background flow. Time goes from left to right. The red regions
  correspond to influx and the blue regions correspond to efflux. In all
  cases, the vesicle reaches a steady state circular shape. (b) The
  reduced area of each of the simulations. The particular snapshots in
  (a) occur at the marks along the curve.}
\end{figure*}
The analysis in \S~\ref{subsec:area_dynamics} shows
that a semipermeable inextensible membrane relaxes to a circle at a rate that is
proportional to $\beta$, regardless of the initial membrane shape (for a fixed length):
In the absence of a background flow, a semipermeable inextensible membrane in free
space is quantified in terms of its area dynamics and transition
to the equilibrium shape, which is a circular membrane with a uniform 
membrane tension $\Lambda = 2\pi^2/L^2$.
This is in stark contrast with the impermeable vesicle relaxing to equilibrium that depends on the vesicle reduced area,  which remains constant for $\beta=0$.
%
Our analysis further shows that the rate of change in area enclosed inside a
semipermeable inextensible membrane is proportional to $\beta$. For this reason,
our numerical results are reported in terms of the scaled time $\beta t$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Planar Shear Flow}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \else
  \includegraphics{figures/shearComposite.pdf}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semipermeable inextensible membrane with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b) The equilibrium membrane shape with varying flow rates and
  membrane permeability. The red regions correspond to influx and the
  blue regions correspond to efflux The boxed membranes correspond to the
  simulation results in part (a).}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Poiseuille Flow}
We consider a semipermeable inextensible membrane in the Poiseuille flow $\uu =
U(1-(y/W)^2,0)$, where $U$ is the maximum velocity at $y=0$. The
non-linear effects of the Poiseuille flow on the migration of an
impermeable vesicle have been well-studied~\cite{kao-bir-mis2009}: At a
given value of the maximum velocity $U$, there exists a critical reduced
volume above which the equilibrium vesicle shape is symmetric (parachute
or bullet shapes) and the vesicle stays at $y=0$. Below this critical
reduced volume the equilibrium vesicle moves off the center and is
asymmetric (tank-treading slipper shape). We validated our numerical
codes by comparing against results in~\citet{kao-bir-mis2009} (see
\S~\ref{subsec:numerical_validation}).

@@

At the beginning of the simulations, a semipermeable vesicle is placed
above the center line $y=0$ and it migrates towards the center line as
in the impermeable case~\cite{dan-vla-mis2009}. 
%
%However,
%for an impermeable vesicle, depending on the initial reduced area and
%the maximum flow velocity $U$, the equilibrium vesicle shape is either
%an axisymmetric parachute or bullet, or an asymmetric tank-treading
%slipper~\cite{kao-bir-mis2009}. 
%
%To validate our methods,
%in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
%for the steady state shape of an impermeable vesicle
%vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
%qualitatively agree, Kaoui et al.~report that the vesicles with
%$\alpha=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
%axisymmetric. In our simulations these vesicles are slightly asymmetric,
%are undergoing a tank treading motion, and we therefore report them as
%asymmetric slippers.
%
Using several initial reduced areas and flow rates, we recreate the
impermeable vesicle phase diagram for semipermeable vesicles with $\beta
= 10^{-3}$. Smaller value of $\beta$ has been used for some parameters but the time 
it takes to reach equilibrium (for some this involves a transient tank-treading
motion along the membrane to gradually slow down in the presence of small $\beta$) is too long in these simulations.

@@

\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/parabolicTensions.tikz}
  \else
  \includegraphics{figures/parabolicTensions.pdf}
  \fi
%  \begin{tikzpicture}[scale=1]
%    \node[inner sep=0pt] at (0,0) {
%    \includegraphics[width=\textwidth]{figures/parabolicTensions.pdf}
%    };
%    \node at (-8.2,1.5) {(a)};
%    \node at (-3.8,1.5) {(b)};
%    \node at (+0.6,1.5) {(c)};
%    \node at (+5.0,1.5) {(d)};
%  \end{tikzpicture}
  \caption{\label{fig:parabolicTensions} The equilibrium tension
  distribution of a semipermeable inextensible membrane with $\beta=10^{-3}$ in a
  parabolic flow (Figure~\ref{fig:parabolicComposite}). The flow
  velocities are (a) 200 $\mu$m/s, (b) 400 $\mu$m/s, (c) 600 $\mu$m/s,
  and (d) 800 $\mu$m/s.}
\end{figure*}
 The membrane tension distribution of the equilibrium shapes are shown
 in Figure~\ref{fig:parabolicTensions} for the flow velocities $U$ from
 $200\;\mu$m/s to $800 \;\mu$m/s. We note that in the Poiseuille flow
 the membrane tension is positive in the front, and negative in the rear
 part of the vesicle. Furthermore we also note that the tension
 magnitude is almost linearly proportional to $U$ as in the linear shear
 flow.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a) Snapshots of four
  semipermeable inextensible membranes with $\beta = 10^{-8}$ submerged
  in a fluid with no background flow. Time goes from left to right. The
  red regions correspond to influx and the blue regions correspond to
  efflux. In all cases, the vesicle reaches a steady state circular
  shape. (b) The reduced area of each of the simulations. The particular
  snapshots in (a) occur at the marks along the curve. Three dimensioned
  times are denoted by the dashed lines.}
\end{figure*}
In the absence of a background flow, a semipermeable inextensible
membrane in free space is quantified in terms of its area dynamics and
transition to the equilibrium shape, which is a circular membrane with a
uniform membrane tension $\Lambda = 2\pi^2/L^2$. This is in stark
contrast with the impermeable vesicle relaxing to equilibrium that
depends on the vesicle reduced area, which remains constant for
$\beta=0$. Our analysis in \S~\ref{subsec:area_dynamics} shows that a
semipermeable inextensible membrane relaxes to a circle at a rate that
is proportional to $\beta$, regardless of the initial membrane shape
(for a fixed length). For this reason, our numerical results are
reported in terms of the scaled time $\beta t$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Planar Shear Flow}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \else
  \includegraphics{figures/shearComposite.pdf}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semipermeable inextensible membrane with $\beta = 10^{-3}$ initialized
  with four different reduced areas in a shear flow with two different
  flow rates. (b) The equilibrium membrane shape with varying flow rates
  and membrane permeability. The red regions correspond to influx and
  the blue regions correspond to efflux. The boxed membranes correspond
  to the simulation results in part (a).}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Poiseuille Flow}
We consider a semipermeable inextensible membrane in the Poiseuille flow
$\uu_\infty(\xx) = U(1-(y/W)^2,0)$, where $U$ is the maximum velocity at
$y=0$. The non-linear effects of the Poiseuille flow on the migration of
an impermeable vesicle have been well-studied~\cite{kao-bir-mis2009}: At
a given value of the maximum velocity $U$, there exists a critical
reduced volume above which the equilibrium vesicle shape is symmetric
(parachute or bullet shapes) and the vesicle stays at $y=0$. Below this
critical reduced volume the equilibrium vesicle moves off the center and
is asymmetric (tank-treading slipper shape). We validated our numerical
codes by comparing against results in~\citet{kao-bir-mis2009} (see
\S~\ref{subsec:numerical_validation}).

@@

At the beginning of the simulations, a semipermeable vesicle is placed
above the center line $y=0$ and it migrates towards the center line as
in the impermeable case~\cite{dan-vla-mis2009}. 
%
%However,
%for an impermeable vesicle, depending on the initial reduced area and
%the maximum flow velocity $U$, the equilibrium vesicle shape is either
%an axisymmetric parachute or bullet, or an asymmetric tank-treading
%slipper~\cite{kao-bir-mis2009}. 
%
%To validate our methods,
%in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
%for the steady state shape of an impermeable vesicle
%vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
%qualitatively agree, Kaoui et al.~report that the vesicles with
%$\alpha=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
%axisymmetric. In our simulations these vesicles are slightly asymmetric,
%are undergoing a tank treading motion, and we therefore report them as
%asymmetric slippers.
%
Using several initial reduced areas and flow rates, we recreate the
impermeable vesicle phase diagram for semipermeable vesicles with $\beta
= 10^{-3}$. Smaller values of $\beta$ have been used for some parameters
but the time it takes to reach equilibrium (for some this involves a
transient tank-treading motion along the membrane to gradually slow down
in the presence of small $\beta$) is too long in these simulations.

@@

\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/parabolicTensions.tikz}
  \else
  \includegraphics{figures/parabolicTensions.pdf}
  \fi
%  \begin{tikzpicture}[scale=1]
%    \node[inner sep=0pt] at (0,0) {
%    \includegraphics[width=\textwidth]{figures/parabolicTensions.pdf}
%    };
%    \node at (-8.2,1.5) {(a)};
%    \node at (-3.8,1.5) {(b)};
%    \node at (+0.6,1.5) {(c)};
%    \node at (+5.0,1.5) {(d)};
%  \end{tikzpicture}
  \caption{\label{fig:parabolicTensions} The equilibrium tension
  distribution of a semipermeable inextensible membrane with
  $\beta=10^{-3}$ in a parabolic flow
  (Figure~\ref{fig:parabolicComposite}). The maximum flow velocities are
  (a) 200 $\mu$m/s, (b) 400 $\mu$m/s, (c) 600 $\mu$m/s, and (d) 800
  $\mu$m/s.}
\end{figure*}
 The membrane tension distribution of the equilibrium shapes are shown
 in Figure~\ref{fig:parabolicTensions} for the flow velocities $U$ from
 $200\;\mu$m/s to $800 \;\mu$m/s. We note that in the Poiseuille flow
 the membrane tension is positive in the front, and negative in the rear
 part of the vesicle. Furthermore we also note that the tension
 magnitude is almost linearly proportional to $U$ as in the linear shear
 flow.
